---
title: "Data Input"
author:
- name: Pascal Maas
  affiliation: Metabolomics Analytical Centre, Leiden, The Netherlands
date: "`r Sys.Date()`"
output: html_document
package: mzQuality
vignette: >
    %\VignetteIndexEntry{Data Input}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mzQuality)
```

## Reading data

mzQuality requires a specific format for the input data. This section will
explain how to prepare your data for analysis. The data should be in a
tab-delimited format containing at least the following columns:

-   `aliquot`: The unique identifier for each sample.
-   `compound`: The name of the compound being measured.
-   `area`: The area under the curve for the compound.
-   `type`: The type of sample (e.g., control, test).
-   `injection_time`: The time at which the sample was injected.
-   `batch`: The batch number for the sample.

Other columns are optional and will be matched to either aliquot, compound,
or both. There are two functions that help you import your data into
mzQuality. To read your data, you can use the `readData` function. This
function will read in your data and check for any missing or incorrect
columns. The `buildExperiment` function will then take the data and create an
experiment object that can be used for analysis.

### (Optional) Adding Concentrations
When calibration lines are present in the experiment, and the concentrations
of spiked compounds is known, mzQuality can calculate the concentration of
these compounds in other samples using (weighted) linear regression. This
is automatically done when the `doAnalysis` function is called. Note that in 
the current version of mzQuality, only one sample type can be used for 
calculating concentrations. For instance, either an Academic Calibration (ACAL)
or Plasma Calibration (CAL) can be used. This will be addressed in a future 
version.

To use and calculate concentrations in mzQuality, ensure that your data file
has the column `concentration` added. This column should contain the known 
concentrations for the compounds in the samples where they were spiked. 
For all other samples, i.e. study samples, the value `NA` should be used.

```{r, echo=FALSE, message=FALSE,  results='asis'}
# Example data
file <- system.file("extdata/example.tsv", package = "mzQuality")
df <- read.delim(file, sep = "\t")
columns <- c(
    "aliquot", "compound", "area", "compound_is", "area_is", 
    "type", "injection_time", "batch", "concentration"
)
caption <- "Example data file with required columns (except concentration)"
cat('<div style="overflow-x: auto;">')
knitr::kable(
    df[8:20, columns],
    caption = caption
)
cat('</div>')
```

Then, run the following code to read your data and build an experiment

```{r, echo=TRUE, message=FALSE, eval = FALSE}
# Path to your file
file <- "path/to/file"

# Read the data, performing basic checks
df <- readData(file)

# Build experiment used throughout mzQuality
exp <- buildExperiment(df)

# Continue analysis, plotting and reporting here.
```


## SessionInfo

```{r, echo=FALSE, message=FALSE}
sessionInfo()
```
