---
title: "Getting Started"
author: "Pascal Maas"
date: "`r Sys.Date()`"
output: html_document
vignette: >
    %\VignetteIndexEntry{Getting Started}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{=html}
<style type="text/css">
div.main-container {
    max-width: 70vw !important;
}
</style>
```

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mzQuality)
```

## Reading data

mzQuality requires a specific format for the input data. This section will
explain how to prepare your data for analysis. The data should be in a
tab-delimited format with the following columns:

-   `aliquot`: The unique identifier for each sample.
-   `compound`: The name of the compound being measured.
-   `area`: The area under the curve for the compound.
-   `type`: The type of sample (e.g., control, test).
-   `injection_time`: The time at which the sample was injected.
-   `batch`: The batch number for the sample.

Other columns are optional and will be matched to either aliquot, compound,
or both. There are two functions that help you import your data into
mzQuality. To read your data, you can use the `readData` function. This
function will read in your data and check for any missing or incorrect
columns. The `buildExperiment` function will then take the data and create an
experiment object that can be used for analysis.

```{r loadingData, echo=FALSE, message=FALSE}
# Example data
file <- system.file("example.tsv", package = "mzQuality")
knitr::kable(head(read.delim(file, sep = "\t")))
```

```{r buildingExperiment, echo=TRUE}
file <- system.file("data.RDS", package = "mzQuality")
experiment <- readRDS(file)
```

## Analysis

Once you have your data in the correct format, you can perform the analysis.
The `doAnalysis` function will take the experiment object and perform the
analysis. This function has several parameters that can be adjusted to
customize the analysis. The following steps are performed in the analysis:

1.  **Calculate Ratios**: Calculates compound ratios. The `calculateRatio`
    function has parameters like `removeOutliers`, `useWithinBatch`, and
    `effectNaAsZero`.

2.  **Batch Correction**: Applies batch correction via `addBatchCorrection`.
    Useful for removing technical variation.

3.  **Background Signals**: Removes background noise using
    `backgroundSignals`.

4.  **Matrix Effect**: Detects and adjusts for matrix effects with
    `matrixEffect`.

5.  **Quality Control**: The `ratioQcSample` function performs QC on
    calculated ratios.

6.  **Type Presence**: Uses `typePresence` to check presence of expected
    sample types.

7.  **Median Sample Area**: Calculates the median area per compound with
    `medianSampleArea`.

8.  **Suggested Internal Standards**: Suggests appropriate internal standards
    using `suggestedInternalStandards`.

9.  **Concentration Calculation**: If concentration data is available,
    `calculateConcentrations` estimates them.

10. **Batch Correction for Concentration**: Adjusts concentration data via
    `addBatchCorrectionAssay`.

11. **Carry Over Effect**: Checks for potential carryover between samples
    using `carryOverEffect`.

12. **Blank Limits**: Detects background noise and sets blank limits with
    `blankLimits`.

```{r analysis, echo=FALSE, eval = FALSE}
experiment <- calculateRatio(experiment) %>%
    addBatchCorrection(
        removeOutliers = removeOutliers,
        useWithinBatch = useWithinBatch
    ) %>%
    backgroundSignals(NaAsZero = effectNaAsZero) %>%
    matrixEffect() %>%
    ratioQcSample() %>%
    typePresence() %>%
    medianSampleArea() %>%
    suggestedInternalStandards(
        removeOutliers = removeOutliers,
        useWithinBatch = useWithinBatch
    )

if ("concentration" %in% assayNames(experiment)) {
    experiment <- experiment %>%
        calculateConcentrations() %>%
        addBatchCorrectionAssay(assay = "concentration") %>%
        carryOverEffect() %>%
        blankLimits()
}
```

All these steps have been combined into the `doAnalysis` function for
convenience. The `doAnalysis` function takes several parameters that control
how the analysis is performed. The following parameters are available:

```{r analysisPipeline, echo=TRUE}
# Combine steps into a single wrapper function
experiment <- doAnalysis(experiment, doAll = TRUE)
```

Explain each step, how it's affected by the parameters, what the output is,
how to interpret it.

## Plotting

Explain the different plots, how to interpret them, what the parameters are,
how to change them.

## Exporting

Explain how to export the results, what formats are available, how to
customize the output.

## Example

Full example from start to finish.

## SessionInfo

```{r}
sessionInfo()
```
