---
title: "mzQuality"
author:
- name: Pascal Maas
  affiliation: Metabolomics Analytical Centre, Leiden, The Netherlands
date: "`r Sys.Date()`"
output: html_document
package: mzQuality
vignette: >
    %\VignetteIndexEntry{mzQuality}
    %\VignetteEncoding{UTF-8}
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
div.main-container {
    max-width: 70vw !important;
}
</style>
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>"
)
```

# Introduction

mzQuality is a R package designed to facilitate the analysis of targeted
mass spectrometry data after peak integration. It provides a
comprehensive set of tools for data import, quality control, and
statistical analysis. 

## Why mzQuality

Quality Control is an important step in the analysis of mass
spectrometry data. Existing tools like `phenomis` and `qmtools` are
designed with access to a phenotype in mind. However, lab technicians
and researchers often (deliberately) do not have access to this
information to remain unbiased. mzQuality was developed to perform
outlier detection, batch correction and other quality control steps
without the need for defined phenotypes. It follows the recommendations
of the mQACC for quality control by leveraging on proper batch design
and the addition of various quality control samples.

Additionally, mzQuality is capable of recommending different internal
standards by exhaustively calculating the Relative Standard Deviation of
QC samples (RSDQC) of batch-corrected ratio's. Here, the internal
standard that results in the lowest RSDQC will be recommended.
Furthermore, by supplying calibration line samples and known
concentrations for spiked compounds, mzQuality is able to calculate
absolute concentrations, enabling downstream modelling of metabolic
pathways and between-cohort analyses.

With lab technicians and researchers in mind, mzQuality is designed to
be a user-friendly R package that can be used by anyone with basic
knowledge of R. As such, it features import of data from a variety of
formats, including a generalized tab-delimited format and Sciex OS text
exports. Furthermore, a Shiny application is in the works that enables
all features without the need for programming.

Taken together, mzQuality is a full-fledged tool for quality control for
metabolomics studies.

# Other Quality Control initiatives
Since Quality Control is such a vital part of any metabolomics experiment,
There are other initiatives and R packages that focus on this topic.

## Related R Packages

-   [phenomis](https://bioconductor.org/packages/release/bioc/html/phenomis.html):
    Provides tools for post-processing like quality control and normalization,
    as well as univariate statistics. 
-   [qmtools](https://bioconductor.org/packages/release/bioc/html/qmtools.html):
    Contains functions for processing quantative metabolomics data, including
    imputation, normalization and filtering. 
-   [MatrixQCvis](https://bioconductor.org/packages/release/bioc/html/MatrixQCvis.html):
    Features interactive visualization of data quality metrics for various
    -omics datasets. Also uses the SummarizedExperiment object, enabling 
    integration in current workflows.

## Consortia & Organizations

-   [mQACC](https://www.mqacc.org/): The Metabolomics Quality Assurance and 
Control Consortium (mQACC) is an international consortium of metabolomics 
researchers that aims to improve the quality of metabolomics data.
-   [HUPO](https://www.hupo.org): The Human Proteome Organization (HUPO) is 
    an international organization that aims to promote the study of the human 
    proteome. They have a [Quality Control working group](https://www.hupo.org/Quality-Control) 
    that focuses on quality control in proteomics.

# Using mzQuality

In this document, we demonstrate a typical workflow of the capabilities
of mzQuality. We'll start by loading mzQuality and its functions. Next,
we'll discuss reading and converting data so it can be used throughout
the package. We'll continue with a showcase of the various plots and
their interpretation. Lastly, an export will be made that contains the
mentioned plots, tab-delimited text files for programmatic access and a
human-friendly Excel export for easy interpretation.

## Loading mzQuality

By running the following code, we load mzQuality and all its
functionality.

```{r library, message = FALSE}
library(mzQuality)
```

## Reading Data

To showcase the capabilities of mzQuality, various example files have
been added to the package. Here, we will use a tab-delimited file as an
example. The file contains the following columns:

\<description of each of the columns\>

```{r reading}
# Example data

file <- system.file("extdata/example.tsv", package = "mzQuality")
df <- readData(file)

# Show the data
DT::datatable(df)
```

Internally, mzQuality uses Bioconductors' *SummarizedExperiment* object
to store the data. The function `buildExperiment` allows you to create a
*SummarizedExperiment* object from a data frame by specifying the
following columns:

-   `compoundColumn`: The column containing the compound names
-   `aliquotColumn`: The column containing the sample names
-   `primaryAssay`: The column containing the primary assay data
-   `secondaryAssay`: The column containing the secondary assay data
-   `typeColumn`: The column containing the sample types
-   `qc`: The column containing the QC sample types
-   `secondaryIndex`: The column containing the secondary index (e.g.
    Internal Standard)

Other columns can be used to store additional information. These columns
will be automatically added to the correct slot of the experiment. As an
example, the optional column `concentration` is added containing known
concentrations of spiked compounds in calibration lines.

Using the provided `compoundColumn` and `aliquotColumn`, the
`buildExperiment` function is able to deduce which values should belong
to the `rowData` and `colData` slots of the *SummarizedExperiment*
object. The `primaryAssay` and `secondaryAssay` columns are used as the
compound area and Internal Standard area, respectively. Note that if the
`secondaryAssay` column is not provided, its value will be set to `1`,
effectively negating the Internal Standard. This is useful for when you
data set does not contain Internal Standards.

The `typeColumn` is used to determine the sample types, which are used
to calculate metrics and determine outliers. The `qc` column is used to
determine which samples are Quality Control samples. The
`secondaryIndex` column is used to determine the Internal Standard for
each compound. The `buildExperiment` function will automatically
calculate the compound / Internal Standard ratio for each sample and
store it in the `ratio` assay.

```{r experiment}
exp <- buildExperiment(
    df = df, 
    compoundColumn = "compound", 
    aliquotColumn = "aliquot", 
    primaryAssay = "area",
    secondaryAssay = "area_is",
    typeColumn = "type",
    qc = "SQC",
    secondaryIndex = "compound_is"
)
```

Alternatively, you can supply an already built *SummarizedExperiment*
object to mzQuality. This is common when R-based metabolomics pipelines
are used like *xcms*. To ensure proper functionality, the names in the
experiment need to be converted to the ones mzQuality uses. This can be
done with the `convertExperiment` function:

```{r convertExperiment}
exp <- convertExperiment(
    exp = exp,
    primaryAssay = "area",
    secondaryAssay = "area_is",
    istd = "compound_is",
    type = "type",
    batch = "batch"
)
```

### Inspecting the Experiment

To get an idea of how the data is structured within the
*SummarizedExperiment* object, we can use the `rowData`, `colData`, and
`assays` functions. The `rowData` function returns the data frame
containing the compound names and compound-related information, such as
assigned Internal Standards.

```{r}
DT::datatable(as.data.frame(rowData(exp)))
```

The `colData` function returns the data frame containing the sample
names and sample-related information, such as sample types and batches.

```{r}
DT::datatable(as.data.frame(colData(exp)))
```

Finally, the `assays` function returns the names of the assays (data)
associated with the compounds and samples. Examples of this are peak
areas, ratios, and concentrations.

```{r}
assays(exp)
```

## Performing analyses

The package tests samples for outliers, specifically for (Pooled)
Quality Control (QC) samples using their Compound / Internal Standard
ratio. Furthermore, Study Samples are tested for mis-injections using
their Internal Standard areas. Doing so, mzQuality recommends which
samples are suitable for reporting & further analysis.

```{r analysis, echo = TRUE, message = FALSE}
exp <- doAnalysis(
    exp = exp, 
    removeOutliers = TRUE, 
    useWithinBatch = TRUE, 
    removeBadCompounds = TRUE,
    qcPercentage = 80,
    backgroundPercentage = 40,
    nonReportableRSD = 30
)
```

## Inspecting Results

Once an initial analysis has been performed, mzQuality provides various
functions for inspecting the results. Here, the following plots are
provided:

-   `aliquotPlot`: A plot showing the Compound / Internal Standard ratio
    for each sample, colored by batch and type.
-   `compoundPlot`: A plot showing the Compound / Internal Standard
    ratio for each compound, colored by batch and type.
-   `pcaPlot`: A PCA plot showing the Compound / Internal Standard ratio
    for each sample, colored by batch and type.
-   `qcPlot`: A violin plot showing the distribution of the Compound /
    Internal Standard ratio for each sample, colored by batch and type.
-   `concentrationPlot`: A plot showing the linear model of a
    calibration line, with added points for the predicted concentrations
    of the samples of a given compound. This plot is only available for
    compounds with a known concentration.

The colors used for each sample and/or type can be customized by setting
the values in `exp$color`.

### Aliquot Plot

The aliquot plot shows the distribution of the selected assay per
aliquot, while the boxes are colored per sample type. This plot is
especially useful for two scenarios. First, technical replaticates
should have similar peak areas and peak ratios within a batch. When
visualizing batch-corrected values, differences should be minimized
between these replicates, while biological samples should retain
biological differences. Secondly, when internal standards are spiked,
any miss-injections show up when visualizing the peak area of these
standards. This plot allows to inspect both, across batches for types of
interest.

```{r aliquotplot, fig.width=12, fig.height=4}
aliquotPlot(
    exp = exp,
    assay = "ratio",
    batches = exp$batch[1],
    types = exp$type,
    logTransform = FALSE
)
```

#### Facetting plots

All plots can be facetted using the `facetPlot` function. This allows
you to plot more batches or types in a single plot. The function
technically allows to facet by any column present in `colData`, but in
practice the `type` and `batch` column are especially useful to group
together. By setting the `ncol` parameter, you can influence the number
of columns.

```{r facetplot, fig.width=12, fig.height=12}
aliquotPlot(
    exp = exp,
    assay = "ratio",
    batches = exp$batch,
    types = exp$type,
    logTransform = FALSE
) %>%
    facetPlot(
        by = "batch", 
        ncol = 2
    )
```

### Compound Plot

The compound plot is a scatter plot of a selected compound, across the
samples. Here, any assay can be picked to plot on the y-axis, including
if this value needs to be log-transformed. The parameters `batches` and
`types` can be set to display the values of interest, but default to all
types and batches present. Finally, trend lines can be plotted for
sample types of interest, which defaults to the sample type set in
`metadata(exp)$QC`. These and calibration line sample types are
particular useful to detect trends within- or between batches.

```{r compoundplot, fig.width=12, fig.height=4}
compoundPlot(
    exp = exp,
    assay = "ratio",
    batches = exp$batch[1],
    types = exp$type,
    logTransform = TRUE, 
    withinTrend = TRUE, 
    trendTypes = c("SQC", "ACAL")
)
```

### PCA Plot

The Principal Component Analysis (PCA) plot can be used to inspect the
variation within- and between batches. By default only the first and
second Principal Component are shown, but this can be changed by
providing the `pc1` and `pc2` arguments.

In the plot, the samples are colored by their batch and type. The Study
Samples (SAMPLE) are colored by their batch, while the other samples are
colored by their type. This enables inspection of batch effects before
and after correction. Ideally, the Quality Control samples used for
batch correction should be centered as much as possible, and forming a
tight cluster.

The batches of the Study Samples should ideally overlap en encompass the
aforementioned Quality Control samples. To aid in inspection, a 95%
confidence interval can be

```{r pcaplot, fig.width=12, fig.height=4}
pcaPlot(
    exp = exp,
    assay = "ratio_corrected",
    batches = exp$batch,
    types = c("SQC", "LQC", "SAMPLE"),
    logTransform = TRUE
)
```

### QC Distribution Plot

The violin plot can be used to inspect the distribution of sample types.
While similar to the aliquot plot, the violin plot provides a more
detailed insight into the values plotted. Often used for pooled samples,
it can be used for any sample type by setting the `types` argument.

```{r qcplot, fig.width=12, fig.height=4}
violinPlot(
    exp = exp,
    assay = "ratio_corrected",
    batches = exp$batch,
    types = "SQC",
    logTransform = FALSE, 
    addConfidenceInterval = TRUE
)
```

### Concentration Plot

If both calibration lines and known concentrations for spiked compounds
are supplied, mzQuality will calculate the concentration in all other
samples automatically when `doAnalysis` is used. The Concentration Plot
is a scatter plot with a linear model added, based on the calculated
`ratio` and known `concentration`. The measurements of the other samples
are projected on this line, ideally within the range of the calibration
line.

```{r concentrationPlot, fig.width=12, fig.height=4}
concentrationPlot(
    exp = exp,
    compound = 1,
    assay = "ratio_corrected", 
    batches = exp$batch[1],
    types = c("SQC", "ACAL", "SAMPLE")
)
```

## Generating exports

mzQuality provides a function to export the results of the analysis. The
`createReports` function creates a folder containing the results of the
analysis. The results are separated by *Plots* and *Reports*.

```{r export, echo = TRUE, eval = FALSE}
# create a temporary folder
output <- tempdir()

createReports(
    folder = output,
    project = "myProject",
    exp = exp[1:3, ], 
    makeSummaryReport = TRUE, 
    makeCompoundReport = TRUE, 
    backgroundPercent = 40, 
    cautionRSD = 15, 
    nonReportableRSD = 30, 
    assays = c("area")
)

# List the files in the Project folder 
list.files(file.path(output, "myProject"), recursive = TRUE)
```

# Session info

```{r session-info, echo = FALSE}
sessionInfo()
```
