---
title: "Validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE
)
```

```{r setup, echo = FALSE, warning=FALSE}
suppressPackageStartupMessages(library(mzQuality2))
```


This is the validation report of mzQuality that is generated automatically. It functions as a tool to confirm the mzQuality calculations with manually calculated ones. 
In this doucment, the following metrics are defined:

- RSDQC: The Relative Standard Deviation of QC samples. This metric is calculated by dividing the  standard deviation ratio of QC samples by the mean ratio of QC samples. In this process, `NA` values are omitted: `sd(ratios, na.rm = TRUE) / mean(ratios, na.rm = TRUE)`
- RSDQC Corrected: Similar to `RSDQC` but using batch-corrected ratios instead. Batch correction is here defined as a batch-median correction, where the median ratio per batch of (S)QCs are calculated and used as correction factor. 
- Background Signal: The background signal is the percentage of area in blank samples (`BLANK`) compared to the percentage of area in study samples (`SAMPLE`). The calculation for this is: `mean(BLANK, na.rm = TRUE) / median(SAMPLE, na.rm = TRUE) * 100`

Not coincidentally, these metrics are used to determine if a compound should be selected in an mzQuality analysis.
To test these metrics, a set of compounds have been picked that display a variety of values, with some leading to include them, while others should be excluded.
The following compounds have been picked:

```{r data, echo = FALSE}
path <- r"(E:\mzQuality Validatie\data example set.txt)"
df <- read.delim(path)

summary <- df %>%
    distinct(compound, compound_is)

DT::datatable(summary, colnames = c("Compound", "Internal Standard"))
```


```{r metrics, echo = FALSE}
compounds <- length(unique(df$compound))
aliquots <- length(unique(df$aliquot))
batches <- length(unique(df$batch))

uniq_df <- df %>%
    distinct(aliquot, type)

types <- table(uniq_df$type)
sqcs <- types["SQC"]
blanks <- types["BLANK"]
samples <- types["SAMPLE"]

# read dataframe with compounds
```


This dataset contains `r compounds` compounds and `r aliquots` aliquots. Of which `r sqcs` aliquots of type SQC, `r blanks` of type BLANK and `r samples` of type SAMPLE. These aliquots are spread across `r batches` batches. 

```{r, echo = FALSE, message=FALSE}
# do Analysis
exp <- buildExperiment(df) %>%
    doAnalysis(removeOutliers = F)
```


### Metrics

The table below shows the metrics defined above and additional columns to indicate if the compound should be reported. The `Confidence Level` columns classifies these metrics into a single value, which can be one of the following: 

- High: This compound has both an RSDQC below `15` and a Background Signal percentage below `40`
- Caution: This compound has an RSDQC between `15` and `30` and a Background Signal percentage below `40`
- Low SNR: This compound has an RSDQC below `30` and a Background Signal percentage above `40`. 
- Non Reportable: This compound has an RSDQC above `30` and the Background Signal may be any number

Finally, the last column indicates if the compound can be reported, which is either a `Yes` or a `No` value. 

```{r}
metrics <- rowData(exp)

metrics$backgroundSignal <- metrics$backgroundSignal * 100

backgroundPercent <- 40
cautionRSD <- 15
nonReportableRSD <- 30

metrics$confidence <- ifelse(metrics$backgroundSignal <= backgroundPercent & metrics$rsdqcCorrected < cautionRSD, "High", "Caution")
metrics$confidence <- ifelse(metrics$rsdqcCorrected > nonReportableRSD, "Non Reportable", metrics$confidence)
metrics$confidence <- ifelse(metrics$backgroundSignal > backgroundPercent, "Low SNR", metrics$confidence)
metrics$selected <- ifelse(metrics$use, "Yes", "No")

metrics <- metrics[, c("rsdqc", "rsdqcCorrected", "backgroundSignal", "selected", "confidence")]
DT::datatable(as.data.frame(metrics), colnames = c("RSDQC", "RSDQC Corrected", "Background Signal", "Report?", "Confidence Level"))
```

