---
title: "mzQuality Summary report"
output: 
    html_document:
      toc: true
      self_contained: true
params:
  exp: ""
  assays: ""
  plots: ""
  project: "mzQuality"
  vendor: "Unknown"
theme: lumen

---

<style type="text/css">
div.main-container {
  max-width: 80vw !important;
}
</style>


```{r loadDependencies, echo = FALSE, include=FALSE}
library(SummarizedExperiment)
exp <- params$exp
assays <- params$assays
plots <- params$plots
```

### About this Project
This is an mzQuality summary report. It was generated at `r lubridate::now()` using version `r utils::packageVersion("mzQuality")`. Out of `r ncol(exp)` samples, `r if ("use" %in% colnames(exp)) sum(exp$use)` samples were used for this analysis. Samples that were excluded were: `r  if ("use" %in% colnames(exp)) paste(colnames(exp[, !exp$use]), collapse = ",")`. The sample data consisted of `r  if ("batch" %in% colnames(exp)) length(unique(exp$batch))` batches, measured between `r  if ("datetime" %in% colnames(exp)) sort(exp$datetime)[1]` and `r if ("datetime" %in% colnames(exp)) sort(-exp$datetime)[1]`. The batches contained `r if ("batch" %in% colnames(exp)) paste(as.vector(table(exp$batch)), collapse = ",")` samples respectively. 

See below for the sample data:

```{r sampleTable, echo = FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]

x <- vectorDf(colData(exp))
DT::datatable(as.data.frame(x))
```

In total, `r nrow(exp)` compounds were submitted, of which `r if ("use" %in% colnames(rowData(exp))) sum(rowData(exp)$use) else 0` compounds were used. See the table below for compound data.

```{r compoundTable, echo = FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]

x <- vectorDf(rowData(exp))
cols <- sapply(x, function(z) "numeric" %in% class(z))

DT::datatable(as.data.frame(x[, cols]))
```


# Aliquot Plot {.tabset .tabset-fade .tabset-pills}
The Aliquot plot shows the distribution of the assay(s) `r paste(assays, collapse = ", ")` in `r if ("batch" %in% colnames(colData(exp))) length(unique(exp$batch)) else 0` batches. 

```{r,echo=FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
N <- length(unique(exp$batch))
chunk <- 3.5
height <- chunk * N
```

```{r aliquotPlot, echo = FALSE, fig.width=24, warning=FALSE, fig.height=height, out.width='100%', results = "asis", dpi = 100}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
if ("Aliquot" %in% params$plots) {
    for (assay in params$assays) {
        cat(sprintf("## %s \n", assay))

        p <- aliquotPlot(
                exp = exp, 
                types = exp$type, 
                assay = assay) %>% 
            facetPlot(ncol = 2) + 
            ggplot2::ggtitle(assay)
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# PCA Plot {.tabset .tabset-fade .tabset-pills}

```{r pcaPlot, echo = FALSE, fig.width=18, warning=FALSE, fig.height=8, out.width='100%', results = "asis", dpi = 100}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
if ("PCA" %in% params$plots) {
    for (assay in params$assays) {
        cat(sprintf("## %s \n", assay))
        
        p <- pcaPlot(
                exp = exp, 
                assay = assay,
                doLog = TRUE, 
                types = c("SQC", "LQC", "SAMPLE"),
                addConfidenceInterval = TRUE
        ) %>% 
            mzQuality::addLabels() + 
                ggplot2::ggtitle(assay)
        
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# Violin QC Plot {.tabset .tabset-fade .tabset-pills}

```{r violinPlot, echo = FALSE, fig.width=24, warning=FALSE, fig.height=height, out.width='100%', results = "asis", dpi = 100}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
if ("QC" %in% params$plots) {
    for (assay in params$assays) {
        cat(sprintf("## %s \n", assay))
        p <- violinPlot(
              exp = exp, 
              assay = assay,
              doLog = TRUE, 
              withinTrend = FALSE, 
              addConfidenceInterval = TRUE,
              addMedian = TRUE, 
              types = "SQC" 
        )
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```


# Heatmap Plot {.tabset .tabset-fade .tabset-pills}
```{r heatmap, echo = FALSE, fig.width=18, warning=FALSE, fig.height=8, out.width='100%', results = "asis", dpi = 100}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
if ("Heatmap" %in% params$plots) {
    for (assay in params$assays) {
        cat(sprintf("## %s \n", assay))
        
        p <- heatmapPlot(
                exp = exp, 
                assay = assay, 
                method = "static"
        )
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# RSDQC Plot {.tabset .tabset-fade .tabset-pills}
```{r compoundSummary, echo = FALSE}

exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
if ("RSDQC" %in% params$plots) {
    cat(sprintf("## %s \n", assay))
    
    p <- rsdqcPlot(
            exp = exp, 
            method = "static"
    )
    print(p)
    cat('<hr style="border:1px solid gray">')
    
}
```

