---
title: "mzQuality Compound report"
output: 
    html_document:
      toc: true
      self_contained: true
params:
  exp: ""
  assays: ""
  plots: ""
  project: "mzQuality"
  vendor: "Unknown"
theme: lumen

---

<style type="text/css">
div.main-container {
  max-width: 80vw !important;
}
</style>


```{r loadDependencies, echo = FALSE, include=FALSE}
library(SummarizedExperiment)
exp <- params$exp
assays <- params$assays
plots <- params$plots
```

### About this Project
This is an mzQuality summary report. It was generated at `r lubridate::now()` using version `r utils::packageVersion("mzQuality2")`. Out of `r ncol(exp)` samples, `r if ("use" %in% colnames(exp)) sum(exp$use)` samples were used for this analysis. Samples that were excluded were: `r  if ("use" %in% colnames(exp)) paste(colnames(exp[, !exp$use]), collapse = ",")`. The sample data consisted of `r  if ("batch" %in% colnames(exp)) length(unique(exp$batch))` batches, measured between `r  if ("datetime" %in% colnames(exp)) sort(exp$datetime)[1]` and `r if ("datetime" %in% colnames(exp)) sort(-exp$datetime)[1]`. The batches contained `r if ("batch" %in% colnames(exp)) paste(as.vector(table(exp$batch)), collapse = ",")` samples respectively. 

See below for the sample data:

```{r sampleTable, echo = FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
cols <- sapply(colData(exp), function(x) !"matrix" %in% class(x))

x <- colData(exp)[, cols]
DT::datatable(as.data.frame(x))
```

In total, `r nrow(exp)` compounds were submitted, of which `r if ("use" %in% colnames(rowData(exp))) sum(rowData(exp)$use) else 0` compounds were used. See the table below for compound data.

```{r compoundTable, echo = FALSE, error=FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
cols <- sapply(rowData(exp), function(x) !"matrix" %in% class(x))

x <- rowData(exp)[, cols]
cols <- sapply(x, function(z) "numeric" %in% class(z))

DT::datatable(as.data.frame(x[, cols]))
```


```{r plotSettings, echo=FALSE, error=FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]
N <- length(unique(exp$batch))
chunk <- 3.5
height <- chunk * N
heightType <- chunk * length(unique(exp$type))
```



# Compounds Per Batch {.tabset .tabset-fade .tabset-pills}
```{r compound2, echo=FALSE, fig.width=24, warning=FALSE, fig.height=height, out.width='100%', results = "asis", dpi = 100, class.source = 'fold-hide', error=FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]

for (assay in params$assays) {
        cat(sprintf("## %s \n", assay))
    
        p <- compoundPlotNew(
                exp = exp,
                assay = assay
        )
        
        p <- mzQuality2::facetPlot(
          plot = p, 
          by = "batch", 
          ncol = 2
        )

        if (!is.null(p)) {
            print(p)
            cat('<hr style="border:1px solid gray">')
        }
}
```

# Compounds Per Type {.tabset .tabset-fade .tabset-pills}
```{r compoundsPerType, echo=FALSE, fig.width=24, warning=FALSE, fig.height=heightType, out.width='100%', results = "asis", dpi = 100, class.source = 'fold-hide', error=FALSE}
exp <- params$exp
exp <- exp[rowData(exp)$use, exp$use]

for (assay in params$assays) {
        cat(sprintf("## %s \n", assay))
    
        p <- compoundPlotNew(
                exp = exp,
                assay = assay,
                trendTypes = unique(exp$type), 
                withinTrend = TRUE
        )
        
        p <- mzQuality2::facetPlot(
          plot = p, 
          by = "type", 
          ncol = 1
        )

        if (!is.null(p)) {
            print(p)
            cat('<hr style="border:1px solid gray">')
        }
}
```
