---
title: "mzQuality Compound Report"
output: 
    html_document:
      toc: true
      self_contained: true
params:
  exp: 
  assays: 
  plots: 
  project: "mzQuality"
  vendor: "Unknown"
theme: lumen
vignette: >
  %\VignetteIndexEntry{Compound Report Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{=html}
<style type="text/css">
div.main-container {
    max-width: 70vw !important;
}
</style>
```

```{r loadDependencies, echo = FALSE, include=FALSE}
library(mzQuality)

if (!isValidExperiment(params$exp)) {
  # use a default example instead
  exp <- readRDS(system.file("extdata/data.RDS", package = "mzQuality"))
  exp <- doAnalysis(exp)
  
  # Select compound number 2 for this analysis
  exp <- exp[2, ]
  assays <- c("ratio", "ratio_corrected")
} else {
  exp <- params$exp
  assays <- params$assays
}

numOfSamples <- ncol(exp)
numOfCompounds <- nrow(exp)
usedSamples <- 0
usedCompounds <- 0
excludedSamples <- ""
numOfBatches <- 1

columns <- colnames(colData(exp))
if ("use" %in% columns){
  usedSamples <- sum(exp$use)
  excludedSamples <- paste(colnames(exp[, !exp$use]), collapse = ",")
} 

if ("batch" %in% columns) {
  numOfBatches <- length(unique(exp$batch))
}

startDate <- "unknown"
endDate <- "unknown"
if ("datetime" %in% columns) {
  dates <- sort(exp$datetime)
  startDate <- dates[1]
  endDate <- dates[length(dates)]
}

columns <- colnames(rowData(exp))
if ("use" %in% columns){
  usedCompounds <- sum(rowData(exp)$use)
} 
exp <- exp[, exp$use]

chunkSize <- 3.5
height <- chunkSize * numOfBatches
heightType <- chunkSize * length(unique(exp$type))
```


### About this Project
This is an mzQuality summary report. It was generated at `r Sys.time()` 
using version `r utils::packageVersion("mzQuality")`. Out of 
`r numOfSamples` samples, `r usedSamples` samples were used for this analysis. 
Samples that were excluded were: `r excludedSamples`. The sample data consisted 
of `r numOfBatches` batches, measured between `r startDate` and `r endDate`. 

```{r sampleTable, echo = FALSE}

DT::datatable(as.data.frame(cleanDataframe(colData(exp))))
```

In total, `r numOfCompounds` compounds were submitted, of which 
`r usedCompounds` compounds were used. See the table below for compound data.

```{r compoundTable, echo = FALSE, error=FALSE}
DT::datatable(cleanDataframe(rowData(exp), onlyNumeric = TRUE))
```

# Compounds Per Batch {.tabset .tabset-fade .tabset-pills}
```{r compound2, echo=FALSE, fig.width=24, warning=FALSE, fig.height=height, out.width='100%', results = "asis", dpi = 100, class.source = 'fold-hide', error=FALSE}

if (nrow(exp) > 0) {
    for (assay in assays) {
        cat(sprintf("## %s \n", assay))
    
        p <-  compoundPlot(
            exp = exp,
            assay = assay, 
        ) %>%
          facetPlot(
            by = "batch", 
            ncol = 2
        )
        
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# Compounds Per Type {.tabset .tabset-fade .tabset-pills}
```{r compoundsPerType, echo=FALSE, fig.width=24, warning=FALSE, fig.height=heightType, out.width='100%', results = "asis", dpi = 100, class.source = 'fold-hide', error=FALSE}

if (nrow(exp) > 0) {

    for (assay in assays) {
            cat(sprintf("## %s \n", assay))
        
            p <- compoundPlot(
                    exp = exp,
                    assay = assay,
                    trendTypes = unique(exp$type), 
                    withinTrend = TRUE
            ) %>%
              facetPlot(
                by = "type", 
                ncol = 1
            )
    
            print(p)
            cat('<hr style="border:1px solid gray">')
    }
}
```
