---
title: "mzQuality Summary Report"
output: 
    html_document:
      toc: true
      self_contained: true
params:
  exp: 
  assays: 
  plots: 
  project: "mzQuality"
  vendor: "Unknown"
theme: lumen
vignette: >
    %\VignetteIndexEntry{Summary Report Example}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---

```{=html}
<style type="text/css">
div.main-container {
    max-width: 70vw !important;
}
</style>
```


```{r loadDependencies, echo = FALSE, include=FALSE}
library(mzQuality)

if (!isValidExperiment(params$exp)) {
  # use a default example instead
  exp <- readRDS(system.file("extdata/data.RDS", package = "mzQuality"))
  exp <- doAnalysis(exp)
  
  assays <- c("ratio", "ratio_corrected")
  plots <- c("Aliquot", "PCA", "QC", "Heatmap", "RSDQC")
} else {
  exp <- params$exp
  assays <- params$assays
  plots <- params$plots
}

numOfSamples <- ncol(exp)
numOfCompounds <- nrow(exp)
usedSamples <- 0
usedCompounds <- 0
excludedSamples <- ""
numOfBatches <- 1

columns <- colnames(colData(exp))
if ("use" %in% columns){
  usedSamples <- sum(exp$use)
  excludedSamples <- paste(colnames(exp[, !exp$use]), collapse = ",")
} 

if ("batch" %in% columns) {
  numOfBatches <- length(unique(exp$batch))
}

startDate <- "unknown"
endDate <- "unknown"
if ("datetime" %in% columns) {
  dates <- sort(exp$datetime)
  startDate <- dates[1]
  endDate <- dates[length(dates)]
}

columns <- colnames(rowData(exp))
if ("use" %in% columns){
  usedCompounds <- sum(rowData(exp)$use)
} 

exp <- exp[, exp$use]

chunkSize <- 3.5
height <- chunkSize * numOfBatches
heightType <- chunkSize * length(unique(exp$type))
```

### About this Project
This is an mzQuality summary report. It was generated at `r Sys.time()` 
using version `r utils::packageVersion("mzQuality")`. Out of 
`r numOfSamples` samples, `r usedSamples` samples were used for this analysis. 
Samples that were excluded were: `r excludedSamples`. The sample data consisted 
of `r numOfBatches` batches, measured between `r startDate` and `r endDate`. 

See below for the sample data:

```{r sampleTable, echo = FALSE}
x <- cleanDataframe(colData(exp))
DT::datatable(as.data.frame(x))
```

In total, `r nrow(exp)` compounds were submitted, of which `r if ("use" %in% colnames(rowData(exp))) sum(rowData(exp)$use) else 0` compounds were used. See the table below for compound data.

```{r compoundTable, echo = FALSE}
x <- cleanDataframe(rowData(exp), onlyNumeric = TRUE)
DT::datatable(x)
```


# Aliquot Plot {.tabset .tabset-fade .tabset-pills}
The Aliquot plot shows the distribution of the assay(s) `r paste(assays, collapse = ", ")` in `r if ("batch" %in% colnames(colData(exp))) length(unique(exp$batch)) else 1` batch(es). 

```{r aliquotPlot, echo = FALSE, fig.width=24, warning=FALSE, fig.height=height, out.width='100%', results = "asis", dpi = 100}
if ("Aliquot" %in% plots) {
    for (assay in assays) {
        cat(sprintf("## %s \n", assay))

        p <- aliquotPlot(
                exp = exp, 
                types = exp$type, 
                assay = assay) %>% 
            facetPlot(ncol = 2) + 
            ggplot2::ggtitle(assay)
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# PCA Plot {.tabset .tabset-fade .tabset-pills}

```{r pcaPlot, echo = FALSE, fig.width=18, warning=FALSE, fig.height=8, out.width='100%', results = "asis", dpi = 100}
if ("PCA" %in% plots) {
    for (assay in assays) {
        cat(sprintf("## %s \n", assay))
        
        p <- pcaPlot(
                exp = exp, 
                assay = assay,
                logTransform = TRUE, 
                types = c("SQC", "LQC", "SAMPLE"),
                addConfidenceInterval = TRUE, 
                addLabels = TRUE
        ) +
          ggplot2::ggtitle(assay)
        
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# Violin QC Plot {.tabset .tabset-fade .tabset-pills}

```{r violinPlot, echo = FALSE, fig.width=24, warning=FALSE, fig.height=8, out.width='100%', results = "asis", dpi = 100}
if ("QC" %in% plots) {
    for (assay in assays) {
        cat(sprintf("## %s \n", assay))
        p <- violinPlot(
              exp = exp, 
              assay = assay,
              logTransform = TRUE, 
              withinTrend = FALSE, 
              addConfidenceInterval = TRUE,
              addMedian = TRUE, 
              types = "SQC" 
        )
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```


# Heatmap Plot {.tabset .tabset-fade .tabset-pills}
```{r heatmap, echo = FALSE, fig.width=18, warning=FALSE, fig.height=8, out.width='100%', results = "asis", dpi = 100}
if ("Heatmap" %in% plots) {
    for (assay in assays) {
        cat(sprintf("## %s \n", assay))
        
        p <- heatmapPlot(exp = exp, assay = assay)
        print(p)
        cat('<hr style="border:1px solid gray">')
    }
}
```

# RSDQC Plot {.tabset .tabset-fade .tabset-pills}
```{r compoundSummary, echo = FALSE, fig.width=18, warning=FALSE, fig.height=8, out.width='100%', results = "asis", dpi = 100}
if ("RSDQC" %in% plots) {
   for (assay in assays) {
    cat(sprintf("## %s \n", assay))
    
    p <- rsdqcPlot(exp = exp)
    print(p)
    cat('<hr style="border:1px solid gray">')
   }
}
```

